\documentclass[a4paper, 10pt]{article}

% Encoding
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

% Margins
\usepackage[left=2.5cm, right=2.5cm]{geometry}

% Mathematics
\usepackage{amsmath, amssymb, mathtools, amsthm}

% Bibliography (uncomment if needed)
\usepackage{csquotes}
\usepackage[backend=biber, sorting=none, style=nature]{biblatex}
\addbibresource{biblio.bib}

% Personal taste
\usepackage[font=footnotesize]{caption}
\usepackage[parfill]{parskip}


\title{HFD Letter of Intent}
\author{}

\begin{document}

\maketitle
One the strategies biological systems have evolved to interact with the
surrounding environment is locomotion. This ability is obtained by a number of
different means, ranging from fundamental protein complexes (cilia, flagella),
to sophisticated organs driven by complex, sequential motor programs. How these
programs are learned, encoded, and recalled in the brain represent still an
active field of research.

It is a widely accepted hypothesis that the cerebellum plays a pivotal role in motor
control, enabling coordination and fine-tuning during the execution of a motor
program. How this function is implemented in the cerebellar circuitry and what
are the driving theoretical principles driving it are still active fields of
research.

Traditionally, cerebellar computations has been considered segregated from other
cortical and sub-cortical areas. In particular, it was thought that their
function was mainly mediated by the cerebello-thalamo-cortical (CTC) loop
\cite{Horne1995TheRO, Gilman1985TheCA}, where the sensory information fed to the
cerebellum from the cortical sensory areas is processed and then relayed to
obtain a movement fine tuning. This view has been substantially challenged
in the last decades with the discovery of several different pathways strongly
coupling the cerebellar activity with other subcortical regions and, in
particular, the dopaminergic and reward systems \cite{Bostan2010TheBG,
Washburn2024TheCD, Wagner2017CerebellarGC, Ohmae2015ClimbingFE}. This coupling
has been shown to directly translate to the involvement of the cerebellum in
tasks not strictly related to motor tuning \cite{Strick2009CerebellumAN,
Overwalle2014SocialCA, Schmahmann2006CognitionEA}, suggesting that the
distinguishing feature of the cerebellum could be linked not to its specific
motor implications but to a more abstract, functional principle. 

Several computational models have been proposed to investigate cerebellar
computations, focusing on precise functional aspects and leveraging different
theoretical approaches \cite{DeSchepper2021ModelSU, DAngelo2016ModelingTC,
Diedrichsen2019UniversalTO, Houghton2021EditorialIP}. A common view, able to
bridge these results into a unified perspective is still, however, lacking. We aim to
address this question with an integrated approach, combining model selection
and inference theory with experimental validation and neuromorphic applications.

First, we are interested in understanding the equivalence of the considered
models in respect to their emerging behaviour. This question can be framed as a
degeneracy problem [GoncaSnail], where the study of invariant parameter
subspaces can be extended to include differences in model functional forms. To
this end, we will pursue two different approaches. In the first, we will make
use of simulation based inference [Lueckermann, Macke, etc.], recurring to the
latest toolkit developments achieved in our group [GoncaSBI, Jaxley]. This
approach, will not only allow us to explore the equivalence between the
different models, but also to study their respective parameter spaces in terms of
emerging computational properties (e.g., movement robustness, energy
minimization [Jedlicka]) in line with the work presented in [VogelsMeta].
Moreover, we would be interested to explore if these properties can be directly
linked to the ability of the model to show low-dimensional mechanics
[MullerParallel]. To this end, we aim at extending our study to summary
statistics derived from several different dimensionality reduction techniques,
both enstablished and still under development in the neuroscientific literature
[PCA, CCA, dPCA, DMD, SSM, Mante and co]. This will allow us in turn to
understand also the power of these techniques when applied to (simulated)
neuronal data, and to outline with more precision their domain of applicability.

The alternative approach we aim at pursuing to address cerebellar modelling is
closely related to variational inference [Bishop]. This approach has lead to a
number of breakthroughs in parameter optimization, in particular in the deep
learning field [VAE, HVAE, DIFFUSION]. By making use of this framework, we aim
to study how the different models of cerebellar function relate to each other in
terms of a latent, not accessible feature space. Ultimately, analyzing this
space would allow us not only to gain a better understanding of their behaviour,
but to identify which theoretical principles could be underpinning the cerebellar
ability to precisely regulate motor and learning function.

Ultimately, we plan on refining these principles in a unified model, able to
provide falsifiable hypotheses on the single cell, the circuit and the
behavioural level. We also plan on using these principles to design new
solutions for spiking and neuromorphic networks, offering valuable contribution
to the computational motor control field. 


\printbibliography

\end{document}
